Measurement of ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross sections at the LArIAT experiment

### 2018-06-25 Chapter-5/5-Preparatory.tex
s

\subsection{G4Beamline}\label{ch:beamlineComposition}
G4Beamline simulates the beam collision at the LArIAT secondary target, the energy deposited by the particles in the LArIAT beamline detectors, and the action of the LArIAT magnets, effectively accounting for particle transport through the beamline from the LArIAT target until ``Big Disk", a fictional, void detector located just before the LArIAT cryostat. 
 At the moment of this writing, G4Beamline does not simulated the responses of the beamline detectors. It is possible to interrogate the truth level information of the simulated particles in several points of the geometry. In order to ease the handshake between G4Beamline and the DDMC, we ask for the beam composition just after WC4.
Since LArIAT data are taken under different beam conditions, we need to simulate separately the beam composition according to the magnets' settings and the secondary pion beam intensity with G4Beamline. For the pion cross section analysis the relevant beam conditions are  secondary pion beam energy of 64 GeV, negative polarity magnet with current of 100 A and 60 A. For the kaon cross section analysis the relevant beam conditions is a secondary  pion beam energy of 64 GeV, positive polarity magnet with current of 100 A.

\subsubsection{Beam Composition for Negative Pion Cross Section}
Even if pions are by far the biggest beam component in negative polarity runs, the LArIAT tertiary beam is not a pure pion beam. While useful to discriminate between pions, kaons, and (anti)protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis. Thus, we need to assess the contamination from beamline particles other than pions in the event selections used for the pion cross section analysis and correct for this background. The first step of this process is assessing the percentage of electrons and muons in the $\pi/\mu/e$ beamline candidates via the G4Beamline MC, as we deem the percentage of kaons and antiprotons negligible after the mass selection (< 1 \textperthousand). 
Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
Figure \ref{fig:BeamComposition} shows the momentum predictions from G4Beamline overlaid with data for the 60A runs (left) and for the 100A runs (right). The predictions for electrons, muons and pions have been staggered and their sum is area normalized to data. Albeit not perfect, these plots show a reasonable agreement in momentum shapes between data and MC.  We attribute the difference in shape (longer tail in data) to a two approximations performed in the MC. Firstly, G4Beamline lacks the simulation of the WC efficiency which is momentum dependent and leads to enhance the number events in the center of the momentum distribution. Secondly, G4Beamline stop tracking pions and their products if they decay after WC1; in data, pion decays in flight can still create a tigger if the produced muon travels thought the beamline detectors. In the pion cross section analysis, these differences between data and the G4Beamline prediction are accounted for as a systematic uncertainty related to the beam composition (see Section \ref{ch:BKGsubXSPuppa}).
%We attribute  the difference in shape to the lack of simulation of the WC efficiency in the MC which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.

@@

\subsubsection{Beam Composition for Positive Kaon Cross Section}
In the positive polarity runs, the tertiary beam composition is mainly pions and protons. The left side of Figure \ref{fig:BeamCompositionPos} shows the  predictions for the momentum spectra for the 100A positive runs  according to  G4Beamline (solid colors) overlaid with data (black points). 
Since the LArIAT beamline detectors can discriminate between kaons and other particles, we do not rely on the G4Beamline simulation to estimate the beamline contamination in the pool of kaon candidates (as in the case of the pion cross section), but rather we use a data drive approach. 
The basic idea of this data driven approach is to estimate the bleed over from high and low mass peaks under the kaon peak by fitting the tails of the $\pi/\mu/e$ and proton mass distributions, as shown in Figure \ref{fig:BeamCompositionPos} right side. 
Since the shape of the tails is unknown, the estimate is done multiple times varying the range and shape for reasonable functions. 
For example, to estimate the proton content under the kaon peak, we start by fitting the left tail of the proton mass distribution with a gaussian function between 650 $MeV/c^2$ and 750 MeV/c$^2$. % in a reasonable range and with a reasonable function. 
 We extend the fit function under the kaon peak and integrate the extended fit function between 350-650 MeV/c$^2$. We integrate the mass histogram in the same range and calculate the proton contamination as the ratio between the two integrals. We repeat this procedure for several fit shapes (gaussian, linear and exponential functions) and tail ranges. Finally, we calculate the contamination as the weighted average of single estimates, where the weights are calculated to be the $1./|1-\chi^2|$ of the tail fits. The procedure is repeated for lighter particles mass peak independently.
With 12 iterations of this method we find a proton contamination of  5.0 $\pm$ 2.0 \%  and a contamination from the lighter particles of 0.2 $\pm$ 0.5 \% .
The estimate of the proton background is currently not used in the kaon cross section analysis, but it is a fundamental step to retrieve the true kaon cross section which will be implemented in the further development of the analysis.

@@

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Contributions from the Beamline Background}\label{ch:PionXSBkgSub2}
We define beamline background every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
\begin{itemize}
\item[]1) electrons,
\item[]2) muons,
\item[]3) secondaries from pion interactions in beamline elements,
\item[]4) matched pile up events.
\end{itemize}

The first step to quantify the effect of the beamline background on the pion cross section is to estimate what percentage of events used in the cross section calculation is not a primary pion.  We start by noting that the last type of background, the ``matched pile up" events, is a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small. %\textcolor{red}{SHOW VTX distribution in WC2TPC match}
As shown in Section \ref{ch:beamlineComposition}, we use G4Beamline to estimate the percentage of pions, muons and electrons at WC4, obtaining the composition shown in Table \ref{tab:beamline}. The next step is to simulate those pions, muons and electrons from WC4 to the TPC with the DDMC and evaluate their contribution to the cross section. To do so, we start by generating the same number of electrons, muons and pions with the DDMC and we apply the same selection chain applied in data (i.e. track multiplicity rejection, WC2TPC match and shower rejection) on the three samples. The number of events per particle species surviving this selection is shown on table \ref{tab:MCafterCutContaminants}. In order to reproduce the data beamline composition, we weight each event of a given particle species according to the estimated content for that species as found with the G4Beamline simulation. In case of 60A runs, for example, the weights are 0.688 for pions,  0.046 for muons  and 0.266 for electrons.

It should be noted that pions may  interact hadronically in the steel (cryostat wall) or in the non-instrumented argon upstream to the TPC front face while traveling the length of between WC4 and the TPC. Or, they could decay in flight between WC4 and the TPC. One of the interaction or decay products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation. We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4.  We estimate the number of secondaries using the DDMC pion sample.  The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4.  The secondary particles to primary pion ratio is 4.9\% in the 60A sample and 4.3\% in the 100A sample.

@@

We derive an estimate of the energy loss between the beamline momentum measurement and the TPC ($E_{loss}$) from the pion and kaon DDMC samples, since this quantity is not  measurable directly on data. 
The $E_{loss}$ distribution for the 60A  and 100A pion sample is shown in figure \ref{fig:ELoss60A}, left and right respectively. The $E_{loss}$ distribution for the  whole kaon sample is shown in figure \ref{fig:ELossKaons}. A clear double peaked structure is visible. After a long investigation on the origin of this unexpected shape, we determined that its origin is due to the particles either missing or hitting the HALO paddle (see Section \ref{sec:MuRS} for the HALO paddle description): a schematic rendering of this occurrence is  shown in figure \ref{fig:Halo}. The kinematic at WC4 determines the trajectory of a particle and indicates whether or not it will hit the halo paddle, whose positioning and central hole size were not sufficiently well defined when the beamline instrumentation was installed. In Figure \ref{fig:PxVsXTrue} , we plot the true  horizontal component of the momentum $P_x$ versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the -60A MC simulation runs -- analogous plots are obtained with the -100A pion simulation and with the kaon simulation. These distributions can be separated drawing a line in this position-momentum space. 
We use a logistic regression  \cite{agresti2013categorical}  as a classifier to find the best separating line, shown in both plots as the red line. We classify as ``hitting the halo paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02* X - 0.4 < 0 $$ and as ``missing the halo  paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02*X - 0.4 > 0, $$ where the coefficients of the line are empirically found by the logistic regression estimation. Overall, this simple method classifies in the right category (hit or miss) about 86\% of the pion events. In MC, we assign  $E_{loss} = 32 \pm 4 $~MeV for pion events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for pion events classified as ``missing the halo paddle". These values are the average and width of the two Landau distributions underneath the double peaked distribution. We apply the same classifier on data.

@@

\begin{figure}[hbpt]
\centering
\includegraphics[scale=0.5]{Chapter-5/Images/Halo.png}
\caption{Schematic rendering of the particle path between WC4 and the TPC front face. The paddle with the hollow central circle represents the Halo paddle. We illustrate two possible trajectories: in black, a trajectory that miss the paddle and goes through the hole in the Halo, in blue a trajectory that hits the Halo paddle and goes through the scintillation material.}
\label{fig:Halo}
\end{figure}

@@

\begin{equation}  
\theta = \cos^{-1} \Big(\frac{\vec p _{\text{incoming}}  \cdot\vec p _{\text{outgoing}}}{|\vec p _{\text{incoming}}|  |\vec p _{\text{outgoing}}| }\Big).
\end{equation}   
In case of the reaction channel,  the presence of several topologies requires a more complex definition, as shown in figure \ref{fig:scatterPic}.  We define the scattering angle as the biggest of the angles between the incoming hadron and the visible daughters, where the visible daughters are charged particles that travel more than the average pitch length ($\delta X = 47$ mm) in the detector (see panel a); in case all the daughters are invisible, the angle is assigned to be 90 deg (see panel b). We chose this working definition of scattering angle for inelastic scattering keeping in mind how our tracking reconstruction works: the tracking will stop correctly if none of the daughters is visible in the detector and it is likely to stop correctly if multiple visible daughters form an interaction vertex. The only ``dangerous" case is the production of one charged daughter plus neutrals, which we can study with this working definition of scattering angle (see panel c).

Once we fix the scattering angle definition, we can study the effects of the angular resolution on the cross section by plotting the true Geant4 total hadronic cross section for interaction angles greater than a minimum interaction angle. The left side of Figure \ref{fig:trueWithAngles} shows the  true Geant4 cross section  for interaction angles greater than 0 deg (green), 4.5 deg (red) corresponding to the MC angular resolution, 5.0 deg  (blue) corresponding to the data angular resolution, and 9.0 deg (yellow).  A small $0.5 \text{ deg}$ systematic shift between the mean of the data and MC angular resolution is present, which has a negligible impact on the cross section. The right side of Figure \ref{fig:trueWithAngles} shows the ratio between the  true cross section for interaction angles greater than 5 deg and the true interaction cross section for all angles;  the cross section for angles greater than 5$^\circ$ accounts for more than 80\% than the total cross section in every energy bin.
%, which we account for in the context of the MC efficiency correction to the cross section, as presented in \ref{sec:angSys}.

@@

The energy deposited by the hadron from the TPC front face until the $j^{\text{th}}$ slice is the sum of the measured energy deposited in each previous slabs $E_{i}$, i.e.
\begin{equation}
E_{\text{FF-j}} = \sum_{i<j} E_{i}, 
\end{equation}
where $E_{i}$ is measured in each slab as  the product of the stopping power,  $dE/dX_{i}$,  and the track pitch for that point. Since the measurements of the energy deposited in each slab rely on the same global calorimetric procedure and tracking algorithms, we assume conservatively that that the measurements of $E_{i}$ are not independent from one another; thus, the uncertainty on $E_{\text{FF-j}}$ becomes
\begin{equation}
\delta E_{\text{FF-j}} =  \sum_{i<j}  \delta E_{i} = (j-1) \delta E_{i}, 
\end{equation}
where $\delta E_{i}$ is the uncertainty on the energy loss in one slab of argon.

>>>>

\subsection{G4Beamline}\label{ch:beamlineComposition}
G4Beamline simulates the beam collision at the LArIAT secondary target, the energy deposited by the particles in the LArIAT beamline detectors, and the action of the LArIAT magnets, effectively accounting for particle transport through the beamline from the LArIAT target until ``Big Disk", a fictional, void detector located just before the LArIAT cryostat. 
 At the moment of this writing, G4Beamline does not simulate the responses of the beamline detectors. It is possible to interrogate the truth level information of the simulated particles in several points of the geometry. In order to ease the handshake between G4Beamline and the DDMC, we ask for the beam composition just after WC4.
Since LArIAT data are taken under different beam conditions, we need to simulate separately the beam composition according to the magnets' settings and the secondary pion beam intensity with G4Beamline. For the pion cross section analysis the relevant beam conditions are  secondary pion beam energy of 64 GeV, negative polarity magnet with current of 100 A and 60 A. For the kaon cross section analysis the relevant beam conditions is a secondary  pion beam energy of 64 GeV, positive polarity magnet with current of 100 A.

\subsubsection{Beam Composition for Negative Pion Cross Section}
Even if pions are by far the biggest beam component in negative polarity runs, the LArIAT tertiary beam is not a pure pion beam. While useful to discriminate between pions, kaons, and (anti)protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis. Thus, we need to assess the contamination from beamline particles other than pions in the event selections used for the pion cross section analysis and correct for this background. The first step of this process is assessing the percentage of electrons and muons in the $\pi/\mu/e$ beamline candidates via the G4Beamline MC, as we deem the percentage of kaons and antiprotons negligible after the mass selection ($<$ 1 \textperthousand). 
Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
Figure \ref{fig:BeamComposition} shows the momentum predictions from G4Beamline overlaid with data for the 60A runs (left) and for the 100A runs (right). The predictions for electrons, muons and pions have been staggered and their sum is area normalized to data. Albeit not perfect, these plots show a reasonable agreement in momentum shapes between data and MC.  We attribute the difference in shape (longer tail in data) to a two approximations performed in the MC. Firstly, G4Beamline lacks the simulation of the WC efficiency which is momentum dependent and leads to enhance the number events in the center of the momentum distribution. Secondly, G4Beamline stop tracking pions and their products if they decay after WC1; in data, pion decays in flight can still create a tigger if the produced muon travels thought the beamline detectors. In the pion cross section analysis, these differences between data and the G4Beamline prediction are accounted for as a systematic uncertainty related to the beam composition (see Section \ref{ch:BKGsubXSPuppa}).
%We attribute  the difference in shape to the lack of simulation of the WC efficiency in the MC which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.

@@

\subsubsection{Beam Composition for Positive Kaon Cross Section}\label{ch:protons}
In the positive polarity runs, the tertiary beam composition is mainly pions and protons. The left side of Figure \ref{fig:BeamCompositionPos} shows the  predictions for the momentum spectra for the 100A positive runs  according to  G4Beamline (solid colors) overlaid with data (black points). 
Since the LArIAT beamline detectors can discriminate between kaons and other particles, we do not rely on the G4Beamline simulation to estimate the beamline contamination in the pool of kaon candidates (as in the case of the pion cross section), but rather we use a data drive approach. 
The basic idea of this data driven approach is to estimate the bleed over from high and low mass peaks under the kaon peak by fitting the tails of the $\pi/\mu/e$ and proton mass distributions, as shown in Figure \ref{fig:BeamCompositionPos} right side. 
Since the shape of the tails is unknown, the estimate is done multiple times varying the range and shape for reasonable functions. 
For example, to estimate the proton content under the kaon peak, we start by fitting the left tail of the proton mass distribution with a gaussian function between 650 $MeV/c^2$ and 750 MeV/c$^2$. % in a reasonable range and with a reasonable function. 
 We extend the fit function under the kaon peak and integrate the extended fit function between 350-650 MeV/c$^2$. We integrate the mass histogram in the same range and calculate the proton contamination as the ratio between the two integrals. We repeat this procedure for several fit shapes (gaussian, linear and exponential functions) and tail ranges. Finally, we calculate the contamination as the weighted average of single estimates, where the weights are calculated to be the $1./|1-\chi^2|$ of the tail fits. The procedure is repeated for lighter particles mass peak independently.
With 12 iterations of this method we find a proton contamination of  5.0 $\pm$ 2.0 \%  and a contamination from the lighter particles of 0.2 $\pm$ 0.5 \%.
The estimate of the proton background is currently not used in the kaon cross section analysis, but it is a fundamental step to retrieve the true kaon cross section which will be implemented in the further development of the analysis.

@@

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Contributions from the Beamline Background}\label{ch:PionXSBkgSub2}
We define ``beamline background" as every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
\begin{itemize}
\item[]1) electrons,
\item[]2) muons,
\item[]3) secondaries from pion interactions in beamline elements,
\item[]4) matched pile up events.
\end{itemize}

The first step to quantify the effect of the beamline background on the pion cross section is to estimate what percentage of events used in the cross section calculation is not a primary pion.  We start by noting that the last type of background, the ``matched pile up" events, is a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small. %\textcolor{red}{SHOW VTX distribution in WC2TPC match}
As shown in Section \ref{ch:beamlineComposition}, we use G4Beamline to estimate the percentage of pions, muons and electrons at WC4, obtaining the composition shown in Table \ref{tab:beamline}. The next step is to simulate those pions, muons and electrons from WC4 to the TPC with the DDMC and evaluate their contribution to the cross section. To do so, we start by generating the same number of electrons, muons and pions with the DDMC and we apply the same selection chain applied in data (i.e. track multiplicity rejection, WC2TPC match and shower rejection) on the three samples. The number of events per particle species surviving this selection is shown on Table \ref{tab:MCafterCutContaminants}. In order to reproduce the data beamline composition, we weight each event of a given particle species according to the estimated content for that species as found with the G4Beamline simulation. In case of 60A runs, for example, the weights are 0.688 for pions,  0.046 for muons  and 0.266 for electrons.

It should be noted that pions may  interact hadronically in the steel (cryostat wall) or in the non-instrumented argon upstream to the TPC front face while traveling the length of between WC4 and the TPC. Or, they could decay in flight between WC4 and the TPC. One of the interaction or decay products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation. We call these occurrences ``secondaries" from pion events, with a terminology inspired by Geant4.  We estimate the number of secondaries using the DDMC pion sample.  The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4.  The secondary particles to primary pion ratio is 4.9\% in the 60A sample and 4.3\% in the 100A sample.

@@

We derive an estimate of the energy loss between the beamline momentum measurement and the TPC ($E_{loss}$) from the pion and kaon DDMC samples, since this quantity is not  measurable directly on data. 
The $E_{loss}$ distribution for the 60A  and 100A pion sample is shown in figure \ref{fig:ELoss60A}, left and right respectively. The $E_{loss}$ distribution for the  whole kaon sample is shown in figure \ref{fig:ELossKaons}. A clear double peaked structure is visible. After a long investigation on the origin of this unexpected shape, we determined that its origin is due to the particles either missing or hitting the HALO paddle (see Section \ref{sec:MuRS} for the HALO paddle description): a schematic rendering of this occurrence is  shown in figure \ref{fig:Halo}. The kinematic at WC4 determines the trajectory of a particle and indicates whether or not it will hit the halo paddle, whose positioning and central hole size were not sufficiently well defined when the beamline instrumentation was installed. In Figure \ref{fig:PxVsXTrue} , we plot the true  horizontal component of the momentum $P_x$ versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the -60A MC simulation runs -- analogous plots are obtained with the -100A pion simulation and with the kaon simulation. These distributions can be separated by a line in this position-momentum space. 
We use a logistic regression  \cite{agresti2013categorical}  as a classifier to find the best separating line, shown in both plots as the red line. We classify as ``hitting the halo paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02* X - 0.4 < 0 $$ and as ``missing the halo  paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02*X - 0.4 > 0, $$ where the coefficients of the line are empirically found by the logistic regression estimation. Overall, this simple method classifies in the right category (hit or miss) about 86\% of the pion events. In MC, we assign  $E_{loss} = 32 \pm 4 $~MeV for pion events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for pion events classified as ``missing the halo paddle". These values are the average and width of the two Landau distributions underneath the double peaked distribution. We apply the same classifier on data.

@@

\begin{figure}[hbpt]
\centering
\includegraphics[scale=0.5]{Chapter-5/Images/Halo.png}
\caption{Schematic rendering of the particle path between WC4 and the TPC front face. The paddle with the hollow central circle represents the Halo paddle. We illustrate two possible trajectories: in black, a trajectory that misses the paddle and goes through the hole in the Halo, in blue a trajectory that hits the Halo paddle and goes through the scintillation material.}
\label{fig:Halo}
\end{figure}

@@

\begin{equation}  
\theta = \cos^{-1} \Big(\frac{\vec p _{\text{incoming}}  \cdot\vec p _{\text{outgoing}}}{|\vec p _{\text{incoming}}|  |\vec p _{\text{outgoing}}| }\Big).
\end{equation}   
In case of the reaction channel,  the presence of several topologies requires a more complex definition, as shown in Figure \ref{fig:scatterPic}.  We define the scattering angle as the biggest of the angles between the incoming hadron and the visible daughters, where the visible daughters are charged particles that travel more than the average pitch length ($\delta X = 47$ mm) in the detector (see panel a); in case all the daughters are invisible, the angle is assigned to be 90 deg (see panel b). We chose this working definition of scattering angle for inelastic scattering keeping in mind how our tracking reconstruction works: the tracking will stop correctly if none of the daughters are visible in the detector and it is likely to stop correctly if multiple visible daughters form an interaction vertex. The only ``dangerous" case is the production of one charged daughter plus neutrals, which we can study with this working definition of scattering angle (see panel c).

Once we fix the scattering angle definition, we can study the effects of the angular resolution on the cross section by plotting the true Geant4 total hadronic cross section for interaction angles greater than a minimum interaction angle. The left side of Figure \ref{fig:trueWithAngles} shows the  true Geant4 cross section  for interaction angles greater than 0 deg (green), 4.5 deg (red) corresponding to the MC angular resolution, 5.0 deg  (blue) corresponding to the data angular resolution, and 9.0 deg (yellow).  As reported in Table \ref{tab:AngRes}, we find a $0.5 \text{ deg}$ difference between the mean of the data and MC angular resolution, which has a negligible impact on the cross section. The right side of Figure \ref{fig:trueWithAngles} shows the ratio between the  true cross section for interaction angles greater than 5 deg and the true interaction cross section for all angles;  the cross section for angles greater than 5$^\circ$ accounts for more than 80\% than the total cross section in every energy bin.
%, which we account for in the context of the MC efficiency correction to the cross section, as presented in \ref{sec:angSys}.

@@

The energy deposited by the hadron from the TPC front face until the $j^{\text{th}}$ slice is the sum of the measured energy deposited in each previous slabs $E_{i}$, i.e.
\begin{equation}
E_{\text{FF-j}} = \sum_{i<j} E_{i}, 
\end{equation}
where $E_{i}$ is measured in each slab as  the product of the stopping power,  $dE/dX_{i}$,  and the track pitch for that point. Since the measurements of the energy deposited in each slab rely on the same global calorimetric procedure and tracking algorithms, we assume conservatively that the measurements of $E_{i}$ are not independent from one another; thus, the uncertainty on $E_{\text{FF-j}}$ becomes
\begin{equation}
\delta E_{\text{FF-j}} =  \sum_{i<j}  \delta E_{i} = (j-1) \delta E_{i}, 
\end{equation}
where $\delta E_{i}$ is the uncertainty on the energy loss in one slab of argon.
